---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# PCCTC Tidy Tuesday event

This is a fun little PCCTC data science team exercise in which we all took 90 minutes to tackle a previous Tidy Tuesday challenge.
Here's my entry that I stitched together in that rush.

---

Packages.

```{r, message=FALSE}
library(tidyverse)
```

Data.

```{r}
cheeses <- read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv'
)
```

From the tidy tuesday readme:

> 248 cheeses have listed fat content. Is there a relationship between fat content and cheese type? What about texture, flavor, or aroma?

Let's start by looking at the fat content column, which needs some help to become numeric.

```{r}
# needs repair to be numeric
cheeses |>
  count(fat_content)

# replace ranges with mean and strip %
cheeses <- cheeses |>
  mutate(
    fat_content_numeric = case_when(
      str_detect(fat_content, "^[0-9]+-[0-9]+%$") ~ {
        nums <- str_match(fat_content, "^([0-9]+)-([0-9]+)%$")
        rowMeans(cbind(as.numeric(nums[,2]), as.numeric(nums[,3])), na.rm = TRUE)
      },
      TRUE ~ as.numeric(str_extract(fat_content, "[0-9]+(\\.[0-9]+)?"))
    )
  )

# check that it worked
cheeses |>
  select(fat_content, fat_content_numeric) |>
  distinct() |>
  arrange(fat_content_numeric) |> 
  head(20)
```

Now cheese type.

```{r}
# cheese name is not helpful: each row is unique
cheeses |> 
  count(n_distinct(cheese))

# subcategorizing by animal source could be interesting
cheeses |> 
  count(milk, sort = TRUE)
```

Any trend by source?

```{r}
cheeses |> 
  filter(
    milk %in% c("cow", "goat", "sheep"),
    !is.na(fat_content_numeric)
  ) |> 
  ggplot(aes(x = milk, y = fat_content_numeric)) +
  geom_boxplot()
```

Moving on to cheese type.

```{r}
# these are silly, just make hard or soft
cheeses |> 
  count(type, sort = TRUE)

cheeses <- cheeses |>
  mutate(
    type_category = case_when(
      str_detect(type, regex("hard|firm", ignore_case = TRUE)) ~ "hard",
      str_detect(type, regex("soft", ignore_case = TRUE)) ~ "soft",
      TRUE ~ NA_character_
    )
  )

cheeses |>
  select(type, type_category) |>
  distinct() |>
  head(20)
```

Any trend by type?

```{r}
cheeses |> 
  filter(
    !is.na(type_category),
    !is.na(fat_content_numeric)
  ) |> 
  ggplot(aes(x = type_category, y = fat_content_numeric)) +
  geom_boxplot()
```
