---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# PCCTC Tidy Tuesday event

This is a fun little PCCTC data science team exercise in which we all took 90 minutes to tackle a previous Tidy Tuesday challenge.
Here's my entry that I stitched together in that rush.

---

Packages.

```{r, message=FALSE}
library(tidyverse)
library(viridis)
library(ggdist)
library(ggbeeswarm)
```

Data.

```{r}
cheeses <- read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv'
)
```

From the tidy tuesday readme:

> 248 cheeses have listed fat content. Is there a relationship between fat content and cheese type? What about texture, flavor, or aroma?

Let's start by looking at the fat content column, which needs some help to become numeric.

```{r}
# needs repair to be numeric
cheeses |>
  count(fat_content)

# replace ranges with mean and strip %
cheeses <- cheeses |>
  mutate(
    fat_content_numeric = case_when(
      str_detect(fat_content, "^[0-9]+-[0-9]+%$") ~ {
        nums <- str_match(fat_content, "^([0-9]+)-([0-9]+)%$")
        rowMeans(cbind(as.numeric(nums[,2]), as.numeric(nums[,3])), na.rm = TRUE)
      },
      TRUE ~ as.numeric(str_extract(fat_content, "[0-9]+(\\.[0-9]+)?"))
    )
  )

# check that it worked
cheeses |>
  select(fat_content, fat_content_numeric) |>
  distinct() |>
  arrange(fat_content_numeric) |> 
  head(20)
```

Now cheese type.

```{r}
# cheese name is not helpful: each row is unique
cheeses |> 
  count(n_distinct(cheese))

# subcategorizing by animal source could be interesting
cheeses |> 
  count(milk, sort = TRUE)
```

Any trend by source?

```{r}
cheeses |> 
  filter(
    milk %in% c("cow", "goat", "sheep"),
    !is.na(fat_content_numeric)
  ) |> 
  ggplot(aes(x = milk, y = fat_content_numeric)) +
  geom_boxplot()
```

Moving on to cheese type.

```{r}
# these are silly, just make hard or soft
cheeses |> 
  count(type, sort = TRUE)

cheeses <- cheeses |>
  mutate(
    type_category = case_when(
      str_detect(type, regex("hard|firm", ignore_case = TRUE)) ~ "hard",
      str_detect(type, regex("soft", ignore_case = TRUE)) ~ "soft",
      TRUE ~ NA_character_
    )
  )

cheeses |>
  select(type, type_category) |>
  distinct() |>
  head(20)
```

Any trend by type?

```{r}
cheeses |> 
  filter(
    !is.na(type_category),
    !is.na(fat_content_numeric)
  ) |> 
  ggplot(aes(x = type_category, y = fat_content_numeric)) +
  geom_boxplot()
```

Inspect unique values for flavor and aroma.
I'm ignoring texture because I believe it to be mostly simplified by the type (hard or soft).

```{r}
cheeses |>
  count(flavor, sort = TRUE) |> 
  print(n = 30)

cheeses |>
  count(aroma, sort = TRUE) |> 
  print(n = 30)

# using key terms from the above output, pick simplifying categories
cheeses <- cheeses %>%
  mutate(
    flavor_aroma = case_when(
      str_detect(tolower(paste(flavor, aroma)), "creamy|mild|pleasant|sweet|buttery|milky") ~ "Creamy/Sweet/Mild",
      str_detect(tolower(paste(flavor, aroma)), "nutty|sharp") ~ "Nutty/Sharp",
      str_detect(tolower(paste(flavor, aroma)), "strong|pungent|spicy|tangy|full-flavored|savory|salty|acidic|smokey") ~ "Strong/Pungent/Spicy",
      str_detect(tolower(paste(flavor, aroma)), "fresh|grassy|herbal|fruity|mushroom|earthy|goaty") ~ "Fresh/Herbal/Fruity",
      TRUE ~ "Other/Uncategorized"
    )
  )

# seems reasonable
cheeses |> 
  count(flavor_aroma)
```

Any trend by flavor and aroma?

```{r}
cheeses |> 
  filter(
    !is.na(flavor_aroma),
    !is.na(fat_content_numeric)
  ) |> 
  ggplot(aes(x = flavor_aroma, y = fat_content_numeric)) +
  geom_boxplot()
```

---

Go for the summarizing plot.
First, subset to complete data.

```{r}
plot_data <- cheeses |> 
  filter(!is.na(type_category), !is.na(flavor_aroma), !is.na(fat_content_numeric))
```

Try a violin with jitter.

```{r}
ggplot(plot_data, aes(x = flavor_aroma, y = fat_content_numeric, fill = flavor_aroma)) +
  geom_violin(trim = FALSE, alpha = 0.7, color = NA) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5, color = "black") +
  facet_wrap(~ type_category, labeller = as_labeller(c(hard = "🪨", soft = "🧽"))) +
  scale_fill_viridis(discrete = TRUE, option = "C") +
  labs(
    title = "Fat Content by Cheese Type and Flavor/Aroma",
    x = NULL,
    y = "Fat Content (%)",
    fill = "Flavor/Aroma"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title.position = "plot", # ensures title is aligned to the plot area
    plot.title = element_text(hjust = 0), # left-aligns the title
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none",
    panel.grid.major.x = element_blank()
  )
```

Maybe a raincloud?

```{r}
ggplot(plot_data, aes(x = flavor_aroma, y = fat_content_numeric, fill = flavor_aroma)) +
  stat_halfeye(
    adjust = .5, width = .6, .width = 0, justification = -.3, point_colour = NA
  ) +
  geom_jitter(aes(color = flavor_aroma), width = 0.15, alpha = 0.5, size = 1.5) +
  facet_wrap(~ type_category) +
  scale_fill_viridis(discrete = TRUE, option = "C") +
  scale_color_viridis(discrete = TRUE, option = "C") +
  labs(
    title = "Fat Content by Cheese Type and Flavor/Aroma",
    x = "Flavor/Aroma Category",
    y = "Fat Content (%)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none"
  )
```

ggbeeswarm?

```{r}
ggplot(plot_data, aes(x = flavor_aroma, y = fat_content_numeric, color = flavor_aroma)) +
  geom_beeswarm(cex = 1.5, size = 2, priority = "density") +
  stat_summary(fun = median, geom = "crossbar", width = 0.5, color = "black", fatten = 2) +
  facet_wrap(~ type_category) +
  scale_color_viridis(discrete = TRUE, option = "C") +
  labs(
    title = "Fat Content by Cheese Type and Flavor/Aroma",
    x = "Flavor/Aroma Category",
    y = "Fat Content (%)",
    color = "Flavor/Aroma"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none"
  )
```

Among these options, I find the violin easiest on my brain to interpret. 
Let's make that prettier.

```{r}
ggplot(plot_data, aes(x = flavor_aroma, y = fat_content_numeric, fill = flavor_aroma)) +
  geom_violin(trim = FALSE, alpha = 0.7, color = NA) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5, color = "black") +
  facet_wrap(~ type_category, labeller = as_labeller(c(hard = "🪨", soft = "🧽"))) +
  scale_fill_viridis(discrete = TRUE, option = "C") +
  labs(
    title = "Fat Content by Cheese Type and Flavor/Aroma",
    x = NULL,
    y = "Fat Content (%)",
    fill = "Flavor/Aroma"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title.position = "plot", # ensures title is aligned to the plot area
    plot.title = element_text(hjust = 0), # left-aligns the title
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none",
    panel.grid.major.x = element_blank()
  )
```
